Wigglyhedra

### 2024-11-07 FPSAC25/wigglyhedraFPSAC.tex
minor corrections
change intro to cite preprint

%% put your English abstract here, or comment this out if you don't have one yet
%% please don't use custom commands in your abstract / resume, as these will be displayed online
%% likewise for citations -- please don't use \cite, and instead write out your citation as something like (author year)
\abstract{Motivated by categorical representation theory, we study the wigglyhedron. Its boundary complex encodes wiggly dissections (sets of pairwise pointed and non-crossing arcs wiggling around $n+2$ points on a line). Its graph is the Hasse diagram of the sublattice of the weak order induced by wiggly permutations (permutations of~$[2n]$ avoiding~$2j-1 \cdots i \cdots 2j$ for~$i < 2j-1$ and~$2j \cdots k \cdots 2j-1$ for~$k > 2j$). It contains copies of all type~$A_n$ Cambrian lattices and Cambrian fans.}

%% put your French abstract here, or comment this out if you don't have one
\resume{Motivés par la théorie catégorique des représentations, nous étudions \linebreak l'onduloèdre. Son complexe de bord encode les dissections ondulantes (ensembles d'arcs ondulant autour de $n+2$ points sur une ligne, deux-à-deux pointés et sans croisement). Son graphe est le diagramme de Hasse du sous-treillis de l'ordre faible induit par les permutations ondulantes (permutations de~$[2n]$ évitant~$2j-1 \cdots i \cdots 2j$ pour~$i < 2j-1$ et~$2j \cdots k \cdots 2j-1$ pour~$k > 2j$). Il contient des copies de tous les treillis Cambriens et associaèdres Cambriens de type~$A_n$.}

@@

Motivated by categorical representation theory, we consider the wiggly complex (\cref{sec:wigglyPseudotriangulations}), whose vertices are arcs wiggling around $n+2$ points on a line, and whose faces are sets of pairwise pointed and non-crossing wiggly arcs.
It is a $(2n-1)$-dimensional manifold whose maximal faces are wiggly pseudotriangulations of a disc.
We show (\cref{sec:bijection}) that the flip graph on wiggly pseudotriangulations is isomorphic to the cover graph of the sublattice of the weak order induced by wiggly permutations (\cref{sec:wigglyPermutations}), the permutations of~$[2n]$ avoiding~$2j-1 \cdots i \cdots 2j$ for~$i < 2j-1$ and~$2j \cdots k \cdots 2j-1$ for~$k > 2j$.
Inspired by similar constructions for finite type cluster complexes~\cite{HohlwegLangeThomas,HohlwegPilaudStella}, and for support $\tau$-tilting complexes on gentle algebras~\cite{PaluPilaudPlamondon-nonkissing}, we then provide geometric realizations of the wiggly complex.
We first construct the wiggly fan (\cref{sec:wigglyFan}), a complete simplicial fan realizing the wiggly complex and supported by the $\b{g}$-vectors of the wiggly arcs.
We then define the wigglyhedron (\cref{sec:wigglyhedron}), a polytope whose normal fan is the wiggly fan and whose support function is given by the incompatibility numbers of wiggly arcs.
Its graph can be oriented to obtain the Hasse diagram of the wiggly lattice, and it contains copies of all type~$A_n$ Cambrian lattices~\cite{Reading-CambrianLattices} as faces (\cref{sec:CambrianConsiderations}).
To conclude, we conjecture that the wiggly complex of any planar point set is polytopal (\cref{sec:planarPointSets}), and we explore representation theoretic aspects of the wiggly complex (\cref{sec:representationTheory}).

@@

\begin{proposition}
\label{prop:hn-filtration}
  Let \(X \in \mathcal{C}_{n+1}\) be a spherical object.
  Consider the set of all indecomposable summands of the nonzero objects \(H^j(X)\).
  Let \(T\) be the corresponding set of wiggly arcs.
  Then
  \begin{compactenum}
  \item 
    \(T\) is a subset of a wiggly pseudotriangulation.
  \item \(T\) contains at most one of the two irrelevant wiggly arcs.
  \end{compactenum}
\end{proposition}
\begin{proposition}
\label{prop:ext1}
Let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to two wiggly arcs \(\alpha\) and \(\alpha'\).
Then the incompatibility degree of \(\alpha\) and \(\alpha'\) is the dimension of \(\Ext_{\mathcal{A}_m}^1(X,X')\).
Thus \(\alpha\) and \(\alpha'\) are compatible (as in~\cref{def:incompatibilityDegree}) if and only if \({\Ext_{\mathcal{A}_m}^1(X, X') = 0}\).
\end{proposition}

>>>>

%% put your English abstract here, or comment this out if you don't have one yet
%% please don't use custom commands in your abstract / resume, as these will be displayed online
%% likewise for citations -- please don't use \cite, and instead write out your citation as something like (author year)
\abstract{
%Motivated by categorical representation theory, we study the wigglyhedron. Its boundary complex encodes wiggly dissections (sets of pairwise pointed and non-crossing arcs wiggling around $n+2$ points on a line). Its graph is the Hasse diagram of the sublattice of the weak order induced by wiggly permutations (permutations of~$[2n]$ avoiding~$2j-1 \cdots i \cdots 2j$ for~$i < 2j-1$ and~$2j \cdots k \cdots 2j-1$ for~$k > 2j$). It contains copies of all type~$A_n$ Cambrian lattices and Cambrian fans.
Motivated by categorical representation theory, we study the wigglyhedron. Its boundary complex encodes wiggly dissections of $n+2$ points on a line. Its skeleton is the Hasse diagram of the sublattice of the weak order induced by wiggly permutations of~$[2n]$. It contains copies of all type~$A_n$ Cambrian lattices and Cambrian~fans.
}

%% put your French abstract here, or comment this out if you don't have one
\resume{
%Motivés par la théorie catégorique des représentations, nous étudions \linebreak l'onduloèdre. Son complexe de bord encode les dissections ondulantes (ensembles d'arcs ondulant autour de $n+2$ points sur une ligne, deux-à-deux pointés et sans croisement). Son graphe est le diagramme de Hasse du sous-treillis de l'ordre faible induit par les permutations ondulantes (permutations de~$[2n]$ évitant~$2j-1 \cdots i \cdots 2j$ pour~$i < 2j-1$ et~$2j \cdots k \cdots 2j-1$ pour~$k > 2j$). Il contient des copies de tous les treillis Cambriens et associaèdres Cambriens de type~$A_n$.
Motivés par la théorie catégorique des représentations, nous étudions \linebreak l'onduloèdre. Son complexe de bord encode les dissections ondulantes de $n+2$ points sur une ligne. Son graphe est le diagramme de Hasse du sous-treillis de l'ordre faible induit par les permutations ondulantes de~$[2n]$. Il contient des copies de tous les treillis Cambriens et associaèdres Cambriens de type~$A_n$.
}

@@

Motivated by categorical representation theory, we consider the wiggly complex (\cref{sec:wigglyPseudotriangulations}), whose vertices are arcs wiggling around $n+2$ points on a line, and whose faces are sets of pairwise pointed and non-crossing wiggly arcs.
It is a $(2n-1)$-dimensional manifold whose maximal faces are wiggly pseudotriangulations.
We show (\cref{sec:bijection}) that the flip graph on wiggly pseudotriangulations is isomorphic to the cover graph of the sublattice of the weak order induced by wiggly permutations (\cref{sec:wigglyPermutations}), the permutations of~$[2n]$ avoiding~$2j-1 \cdots i \cdots 2j$ for~$i < 2j-1$ and~$2j \cdots k \cdots 2j-1$ for~$k > 2j$.

Inspired by similar constructions for finite type cluster complexes~\cite{HohlwegLangeThomas,HohlwegPilaudStella}, and for support $\tau$-tilting complexes on gentle algebras~\cite{PaluPilaudPlamondon-nonkissing}, we then provide geometric realizations of the wiggly complex.
We first construct the wiggly fan (\cref{sec:wigglyFan}), a complete simplicial fan realizing the wiggly complex and supported by the $\b{g}$-vectors of the wiggly arcs.
We then define the wigglyhedron (\cref{sec:wigglyhedron}), a polytope whose normal fan is the wiggly fan and whose support function is given by the incompatibility numbers of wiggly arcs.
Its graph can be oriented to obtain the Hasse diagram of the wiggly lattice, and it contains copies of all type~$A_n$ Cambrian lattices~\cite{Reading-CambrianLattices} as faces (\cref{sec:CambrianConsiderations}).

We conclude with two further topics.
First, we conjecture that the wiggly complex of any planar point set is polytopal (\cref{sec:planarPointSets}), extending the wigglyhedron for aligned points and the pointed pseudotriangulation polytope of~\cite{RoteSantosStreinu-polytope} for points in general position.
We then explore representation theoretic aspects of the wiggly complex (\cref{sec:representationTheory}), interpreting our combinatorial results in terms of operations in a triangulated category.

Due to space limitations, this extended abstract omits many details and proofs of~\cite{BapatPilaud}.

@@

\begin{proposition}
\label{prop:hn-filtration}
  Let \(X \in \mathcal{C}_{n+1}\) be a spherical object.
  Consider the set of all indecomposable summands of the nonzero objects \(H^j(X)\).
  Let \(T\) be the corresponding set of wiggly arcs.
  Then
  \begin{compactenum}
  \item 
    \(T\) is a subset of a wiggly pseudotriangulation.
  \item \(T\) contains at most one of the two irrelevant wiggly arcs.
  \end{compactenum}
\end{proposition}
\begin{proposition}
\label{prop:ext1}
Let \(X\) and \(X'\) be objects in the standard heart \(\mathcal{A}_m\) corresponding to two wiggly arcs \(\alpha\) and \(\alpha'\).
Then the incompatibility degree of \(\alpha\) and \(\alpha'\) is the dimension of \(\Ext_{\mathcal{A}_m}^1(X,X')\).
Thus \(\alpha\) and \(\alpha'\) are compatible (as in~\cref{def:incompatibilityDegree}) if and only if \({\Ext_{\mathcal{A}_m}^1(X, X') = 0}\).
\vincent{Do you mean \cref{def:compatible} here?}
\end{proposition}
